\section{Results}\label{sec:results}
To systematically evaluate the different configurations of each clustering algorithm, the following procedure is followed to extract results for each of the 3 data sets, in order to later perform an analysis those results:

\begin{enumerate}
    \item \textbf{Data Preparation}: The dataset is loaded and the data samples are separated from their labels into separate 2 sets. This way, we perform the clustering analysis in a completely non-supervised way, and we then utilize the labels to extract supervised metrics of the cultering results.

    \item \textbf{Parameter Configuration}: A comprehensive set of values for the algorithm's hyperparameters is defined. These combinations reflect various ways to tune the clustering algorithm.

    \item \textbf{Model Evaluation}: For each parameter combination, the clustering algorithm is applied on the unlabeled data and then evaluated with different metrics. This step yields the following metrics: Adjusted Rand Score (ARI), Normalized Mutual Information (NMI), Davies-Bouldin Index(DBI), Silhouette score, Calinski-Harabasz Score (CHS), and execution time. Together, these metrics (the first 2 supervised, and the rest non-supervised) measure the effectiveness and efficiency of the clustering.

    \item \textbf{Results Compilation}: The performance metrics for each parameter combination are recorded in a structured format. These results are saved as a dataset that summarizes the outcomes of all evaluations, forming a basis for analysis. We save as well the cluster labels of all samples for each clustering algorithm that we run, so we can recover the same clusters in the posterior analysis.

    \item \textbf{Results Analysis}: After results are compiled across all configurations, quantitative and qualitative analysis is performed to identify common trends among the different algorithm configurations for each of the datasets. Additionally, we extract the top performing configuration according to each of the 5 evaluation metrics, in order to study common patterns and visualize the resulting clusters. This analysis helps determine the most reliable and effective parameter settings for accurate and efficient clustering for each dataset.
\end{enumerate}
Due to the vast ammount of information that can be extracted from the results, we will only explicitly showcase the most clarifying plots and results that we have extracted. However, all of the information is available in the \texttt{plots\_and\_tables} folder for each of the datasets, in the \texttt{code} attached to this report.

\textit{Note:} Since all of the tested datasets have high dimensionality, we use Principal Component Analysis (PCA) to extract the 2 principal components in order to generate the clustering scatter plots of the datasets for visualization purposes.

\input{sections/results/kmeans}

\input{sections/results/global_kmeans}

% \input{sections/results/fuzzyCMean.tex}
