\subsection{Datasets}
A brief overview of the three datasets is presented below.

\subsubsection{Hepatitis}
The Hepatitis dataset is a medical dataset aimed at predicting patient survival based on clinical and demographic features. 
This dataset is characterized by its imbalance, as it involves a higher number of survivors compared to non-survivors. It 
is primarily used for binary classification tasks, where the objective is to determine survival outcomes. The features in 
this dataset are diverse, including both numerical and categorical attributes, which provide rich information about patients'
 health status, age, and other clinical measurements. 


\subsubsection{Mushroom}

The Mushroom dataset is designed to classify mushrooms as either edible or poisonous based on their physical 
characteristics. It is a balanced dataset with 8,124 samples, divided into 4,208 edible and 3,916 poisonous mushrooms. 
The dataset comprises 23 categorical features, which capture various traits such as gill spacing, cap shape, and odor. 
These features are primarily ordinal and nominal, requiring appropriate encoding before applying machine learning models. 

\subsubsection{Pen-based}

The Pen-Based dataset is a numerical dataset designed for multiclass classification tasks, specifically recognizing handwritten 
digits from 0 to 9. It consists of 7,494 training samples and 3,498 testing samples, each represented by 16 numerical features 
derived from pen-tip trajectories on a 2D grid, with values ranging from -1 to 1. It had no missing values.
